Mechanisms of AIAN Inequality in Child Welfare Across U.S. States.

### 2020-12-15 cfp_final_report.Rmd
start of day update

# National and state-level marginal probabilities

@@

>>>>

## National and state-level marginal probabilities

Figure \@ref(nat-marginal) shows the marginal probabilities of experiencing a series of child welfare system events for both Native and non-Native children. This plot shows national age-specific event probabilities at 2014-2018 risk levels for AIAN and non-AIAN children for experiencing 1) a screened-in investigation, 2) a confirmed maltreatment case, 3) a foster care placement, and 4) termination of parental rights through the child welfare system. DISCUSSION OF THIS.

```{r nat-marginal, fig.cap = "Marginal probability of child welfare event incidence for AIAN children by age 18, 2014 - 2018 risk levels, US totals"}
##############################################################
###### National age-spec risk
##############################################################

# make 2 x 2 age-spec plots for each outcome

## stack tables for faceting
plot_dat<-inv_nat_tab %>% 
  group_by(race_ethn, age) %>% 
  summarise(qmax = max(q) , qmin=min(q) , q = mean(q),
            c = mean(c), cmax = max(c), cmin = min(c)) %>% 
  ungroup() %>% 
  mutate(var = "Investigation") %>% 
  bind_rows(
    sub_nat_tab %>% 
      group_by(race_ethn, age) %>% 
      summarise(qmax = max(q) , qmin=min(q) , q = mean(q),
                c = mean(c), cmax = max(c), cmin = min(c)) %>% 
      ungroup() %>% 
      mutate(var = "Substantiation")
  ) %>% 
  bind_rows(
    fc_nat_tab %>% 
      group_by(race_ethn, age) %>% 
      summarise(qmax = max(q) , qmin=min(q) , q = mean(q),
                c = mean(c), cmax = max(c), cmin = min(c)) %>% 
      ungroup() %>% 
      mutate(var = "Foster Care")
  ) %>% 
  bind_rows(
    tpr_nat_tab %>% 
      group_by(race_ethn, age) %>% 
      summarise(qmax = max(q) , qmin=min(q) , q = mean(q),
                c = mean(c), cmax = max(c), cmin = min(c)) %>% 
      ungroup() %>% 
      mutate(var = "Termination")
  ) %>% 
  mutate(var = factor(var,
                      levels = c("Investigation",
                                 "Substantiation",
                                 "Foster Care",
                                 "Termination")))

ggplot(plot_dat,
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           color = race_ethn, fill = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) +
  labs(color = "", fill = "") + 
  labs(x = "Age", y = "Probability") + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2") + 
  facet_wrap(~var)

THEN AGE-SPEC GEO FACET

FOR INVESTIGATIONS

```{r nat-inv-age, fig.cap = "Age specific risk of experiencing first CPS investigation by state, 2014 - 2018 risk levels"}
ggplot(investigation_tables %>% 
         group_by(state, race_ethn, age) %>% 
         summarise(qmax = max(q_c), qmin=min(q_c), q = mean(q_c)),
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           color = race_ethn, fill = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) + 
  facet_geo(~state) + 
  labs(color = "", fill = "") + 
  theme_minimal() + 
  labs(x = "Age", y = "Proportion of children") + 
  theme(strip.text.x = element_text(size = 7),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 4)) + 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")

FOR FOSTER CARE

```{r nat-fc-age, fig.cap = "Age specific probability of entering foster care, 2014 - 2018 risk levels"}
ggplot(fc_tables %>% 
         group_by(state, race_ethn, age) %>% 
         summarise(qmax = max(q_c), qmin=min(q_c), q = mean(q_c)),
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           color = race_ethn, fill = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) + 
  facet_geo(~state) + 
  labs(color = "", fill = "") + 
  theme_minimal() + 
  labs(x = "Age", y = "Proportion of children") + 
  theme(strip.text.x = element_text(size = 7),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 4)) + 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")
```

THEN HIGH RISK STATES FOR EACH OUTCOME FOR AGESPEC

THEN CUMULATIVE MAP

```{r cumul-map, fig.cap = "Cumulative risk of child welfare system event for AIAN children by age 18 at 2014 - 2018 risk levels"}
#################################################################
############# MAPS
#################################################################
### Lifetime risk

map_dat<-investigation_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

### xwalk state maps onto abbreviations
us_map<-us_map()

us_map<-us_map %>% 
  left_join(map_dat %>% 
              rename(abbr = state))

inv_map<-ggplot(us_map %>% 
         filter(!(is.na(var))),
       aes(x = x, y = y, group = group,
           fill = c)) + 
  geom_polygon(color = "black") + 
  coord_fixed() +
  theme_void() + 
  scale_fill_gradient2() +
  theme(legend.position = "bottom") +
  labs(fill = "Investigation") +
  guides(fill = guide_colourbar(title.position = "top",
                             label.position = "bottom",
                             title.hjust = 0.5))

map_dat<-subst_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

us_map<-us_map()

us_map<-us_map %>% 
  left_join(map_dat %>% 
              rename(abbr = state))

sub_map<-ggplot(us_map %>% 
                  filter(!(is.na(var))),
                aes(x = x, y = y, group = group,
                    fill = c)) + 
  geom_polygon(color = "black") + 
  coord_fixed() +
  theme_void() + 
  scale_fill_gradient2() +
  theme(legend.position = "bottom") +
  labs(fill = "Substantation") +
  guides(fill = guide_colourbar(title.position = "top",
                                label.position = "bottom",
                                title.hjust = 0.5))

map_dat<-fc_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

us_map<-us_map()

us_map<-us_map %>% 
  left_join(map_dat %>% 
              rename(abbr = state))

fc_map<-ggplot(us_map %>% 
                  filter(!(is.na(var))),
                aes(x = x, y = y, group = group,
                    fill = c)) + 
  geom_polygon(color = "black") + 
  coord_fixed() +
  theme_void() + 
  scale_fill_gradient2() +
  theme(legend.position = "bottom") +
  labs(fill = "Foster Care") +
  guides(fill = guide_colourbar(title.position = "top",
                                label.position = "bottom",
                                title.hjust = 0.5))

map_dat<-tpr_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

us_map<-us_map()

us_map<-us_map %>% 
  left_join(map_dat %>% 
              rename(abbr = state))

tpr_map<-ggplot(us_map %>% 
                 filter(!(is.na(var))),
               aes(x = x, y = y, group = group,
                   fill = c)) + 
  geom_polygon(color = "black") + 
  coord_fixed() +
  theme_void() + 
  scale_fill_gradient2() +
  theme(legend.position = "bottom") +
  labs(fill = "TPR") +
  guides(fill = guide_colourbar(title.position = "top",
                                label.position = "bottom",
                                title.hjust = 0.5))

c_map<-grid.arrange(inv_map, sub_map,
                    fc_map, tpr_map)

THEN INEQ MAP

```{r}
##################
#### #### #### #### 
#### rate ratio for inequality
#### #### #### #### 
### xwalk state maps onto abbreviations
us_map<-us_map()

us_map<-us_map %>% 
  left_join(map_dat %>% 
              rename(abbr = state))

#### DO OTHER MAPS HERE

map_dat<-investigation_tables_c %>% 
  group_by(state, race_ethn) %>% 
  filter(c == min(c)) %>% 
  mutate(var = "Investigation") %>% 
  bind_rows(subst_tables_c %>% 
              group_by(state, race_ethn) %>% 
              filter(c == min(c)) %>% 
              mutate(var = "Substantiation")) %>% 
  bind_rows(fc_tables_c %>% 
              group_by(state, race_ethn) %>% 
              filter(c == min(c)) %>% 
              mutate(var = "Foster Care")) %>% 
  bind_rows(tpr_tables_c %>% 
              group_by(state, race_ethn) %>% 
              filter(c == min(c)) %>% 
              mutate(var = "TPR")) %>% 
  select(-.imp) %>% 
  distinct()

map_dat_wht<-map_dat %>% 
  ungroup() %>% 
  filter(race_ethn == "White") %>% 
  rename(c_white = c_c) %>% 
  select(-race_ethn, -c, -race_ethn)

map_dat<-map_dat %>% 
  filter(race_ethn=="AIAN") %>% 
  left_join(map_dat_wht) %>% 
  mutate(disp = c_c / c_white)

### slice into quantiles
### xwalk state maps onto abbreviations
us_map<-us_map()

us_map<-us_map %>% 
  left_join(map_dat %>% 
              rename(abbr = state)) %>% 
  mutate(var = factor(var,
                      levels = c(
                        "Investigation", 
                        "Substantiation",
                        "Foster Care",
                        "TPR"
                      )))

### cut into an ordered categorical

us_map<-us_map %>% 
  mutate(rate_ratio = 
           case_when(
             disp<0.5 ~ "0-0.5",
             disp<1 ~ "0.5-1",
             disp<1.5 ~ "1-1.5",
             disp<2 ~ "1.5-2",
             disp<3 ~ "2-3",
             disp<4 ~ "3-4",
             disp>=4 ~ "4+",
           ),
         rate_ratio = factor(rate_ratio,
                             levels = c("0-0.5",
                                        "0.5-1",
                                        "1-1.5",
                                        "1.5-2",
                                        "2-3",
                                        "3-4",
                                        "4+")))

ggplot(us_map,
       aes(x = x, y = y, group = group,
           fill = rate_ratio)) + 
  geom_polygon(color = "black") + 
  coord_fixed() +
  facet_wrap(~var) + 
  theme_void() + 
  scale_fill_brewer(palette = "Purples") +
  labs(fill = "AIAN / White\nRate Ratio")
```

@@

```{r}
##############################################################
###### National conditionals
##############################################################

# make 2 x 2 age-spec plots for each conditional

plot_dat<-cond_sub_inv %>% 
  select(.imp, age, race_ethn, cond, var) %>% 
  bind_rows(
    cond_fc_inv %>% 
      select(.imp, age, race_ethn, cond, var)) %>% 
  bind_rows(cond_fc_sub %>% 
              select(.imp, age, race_ethn, cond, var)) %>% 
  bind_rows(
    cond_tpr_fc %>% 
      select(.imp, age, race_ethn, cond, var)
  ) %>% 
  group_by(age, race_ethn, var) %>% 
  summarise(q = mean(cond),
            qmin = min(cond),
            qmax = max(cond))

plot_dat<-plot_dat %>% 
  mutate(var = case_when(
    var == "P(FC|inv)" ~ "2. Foster care after investigation",
    var == "P(FC|sub)" ~ "3. Foster care after substantiation",
    var == "P(sub|inv)" ~ "1. Substantiation after investigation",
    var == "P(TPR|FC)"  ~ "4. Termination after foster care"
  ))

ggplot(plot_dat,
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           fill = race_ethn, color = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) + 
  facet_wrap(~factor(var)) + 
  labs(color = "", fill = "", x = "Age", y = "Conditional probability") + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2") 
```
